Exploring the Role of Feedback in Machine Learning Jupyter Notebooks

### 2024-05-13 report/emse24.tex
feat(report): emse24 updates

\subsubsection{Value check/ Spot checks ($N = 5$)}
% (O56, O60, P64, P67, P114)

@@

O60 &
\begin{lstlisting}
X_pca.head()
\end{lstlisting}\\
\end{tabular}
\caption{Example of outputs used to perform spot checks.}
\label{tab:shape-check}
\end{table}

@@

In \emph{O60}, the analyst verifies that the number of features in the data matches expectations after apply Principal Component Analysis (PCA). This is a critical step for maintaining dimensional consistency in feature-reduced datasets.

% In another case, P64, which is particularly relevant in image processing contexts, the maximum value in the second channel of a 3D numpy array is extracted and verified using np.max(cur[:, :, 1]), ensuring that the data manipulation retains its expected properties.

@@

% Lastly, P114 assesses the correctness of one-hot encoding, a fundamental preprocessing step for categorical data, ensuring that the transformation has been executed correctly as seen with print(onehot_encoded). Each of these checks not only validates individual transformations or predictions but also reinforces the overall reliability and robustness of the machine learning models being developed.

\subsubsection{Model training (O8, O31, O42, P77)}

O8, P77: Periodically print the progress made in training by printing the training loss/accuracy of model.

O31: check the model fits the data without any errors; this is done in the model experimentation phase, where several models are tested and the best one based on the selection criteria is picked

>>>>

\subsubsection{Spot checks ($N = 5$)}
% (O56, O60, P64, P67, P114)

@@

O60 &
\begin{lstlisting}
X_pca.head()
\end{lstlisting}\\

P64 &
\begin{lstlisting}
print(np.max(cur[:, :, 1]))
\end{lstlisting}\\

P114 &
\begin{lstlisting}
print(onehot_encoded)
\end{lstlisting}\\
\end{tabular}
\caption{Example of outputs used to perform spot checks.}
\label{tab:spot-check}
\end{table}

@@

@@

In \emph{O60}, the analyst verifies that the number of features in the data matches expectations after apply Principal Component Analysis (PCA). This is a critical step for maintaining dimensional consistency in feature-reduced datasets. \emph{P64} demonstrates another case which is particularly relevant in image processing contexts. The maximum value in the second channel of a 3D numpy array is extracted and verified, ensuring that the data manipulation retains its expected properties. Lastly, \emph{P114} assesses the correctness of one-hot encoding, a fundamental preprocessing step for categorical data, ensuring that the transformation has been executed correctly.

Each of these checks not only validates individual transformations or predictions but also reinforces the overall reliability and robustness of the machine learning models being developed.

\subsubsection{Model training ($N = 4$)}
% (O8, O31, O42, P77)

O8 &
\begin{lstlisting}
autoencoder.fit(x=X_train, y=X_train, epochs=15, validation_data=[X_test, X_test], callbacks=[keras_utils.TqdmProgressCallback()], verbose=0)
\end{lstlisting}\\

O31 &
\begin{lstlisting}
adaBoost.fit(X_train, y_train)
\end{lstlisting}\\

O42&
\begin{lstlisting}
m_r.best_params_
\end{lstlisting}\\
\end{tabular}
\caption{Example of outputs used to monitor model training.}
\label{tab:model-training}
\end{table}

The ``Model Training'' phase of ML workflows involves various critical activities, each aimed at ensuring effective model performance and robustness.

During training it is common practice to monitor progress periodically, as illustrated by outputs \emph{O8} in Table~\ref{tab:model-training}, where the training loss and accuracy are printed to give real-time feedback on the model's learning efficacy. This feedback allows for adjustments in training parameters or early stopping to prevent overfitting. For instance, in output O8, the autoencoder is trained using a custom callback from \texttt{keras\_utils} to provide a progress update without cluttering the output stream, which is particularly useful in lengthy training sessions. In contrast, output \emph{O31} represents a crucial validation step where the adaBoost model is fitted to the training data to ensure it adheres to the expected data patterns without errors, a typical task in the model experimentation phase where various models are assessed for their suitability. Output \emph{O42} possibly checks for the best parameters found through tuning methods, highlighting the importance of parameter optimisation in achieving optimal model performance.

Each of these outputs underscores different facets of the model training phase, from initialization and real-time monitoring to validation and finalization, illustrating the layered complexity of developing predictive models.
